---
title: "data_comparison"
author: "Hannah Dempsey"
output: github_document
---

# Comparing Sonoran Sites with Portal

Set up:

```{r}
library(tidyverse)
library(portalr)
```

### Biomass Data

#### By species:

Adjusting Portal's biomass for multiple surveys per year:

```{r}
#Loading in data
portal_biomass <- read_csv("../data_clean/portal_biomass.csv")
portal_trapping <- read_csv("../data_clean/portal_trapping.csv")

#calculating number of periods per year
portal_periods <- portal_trapping %>% 
  group_by(year) %>% 
  summarize(periods = unique(period))
portal_periods
 
portal_periods_annual <- portal_periods %>% 
  group_by(year) %>% 
  summarize(n_periods = n())
portal_periods_annual

#dividing biomass by number of periods
adjusted_portal_biomass <- portal_biomass %>% 
  inner_join(portal_periods_annual, join_by(year)) %>% 
  mutate(adjusted_biomass_g = biomass_g / n_periods) %>% 
  select(1:2, 7) %>% 
  #renaming so all column names across sites are the same
  rename(biomass_g = adjusted_biomass_g)
adjusted_portal_biomass
```

Modify portal's species column to match:

```{r}
#load in species data
portal_species <- read_csv("../data_clean/portal_species.csv")

#distinguish between hispidus spp
adjusted_portal_species <- portal_species %>% 
  mutate(scientificname = str_replace(scientificname, "Chaetodipus hispidus", " C.hispidus")) %>% 
  mutate(scientificname = str_replace(scientificname, "Sigmodon hispidus", " S.hispidus")) %>% 
  separate(scientificname, c("genus", "species_name"), sep = " ") %>% 
  select(1, 3)
adjusted_portal_species

#join to biomass dataframe
adjusted_portal_biomass <- adjusted_portal_biomass %>% 
  left_join(adjusted_portal_species, join_by(species)) %>% 
  select(1, 4, 3) %>% 
  rename(species = species_name)
adjusted_portal_biomass
```

Add site column to all dataframes:

```{r}
#loading in the rest of the data
saguaro_biomass <- read_csv("../data_clean/saguaro_biomass.csv")
orpi_biomass <- read_csv("../data_clean/orpi_biomass.csv")

#adding site column to each dataframe
saguaro_biomass <- saguaro_biomass %>% 
  mutate(site = "Saguaro")
saguaro_biomass

orpi_biomass <- orpi_biomass %>% 
   mutate(site = "Organ Pipe")
orpi_biomass

adjusted_portal_biomass <- adjusted_portal_biomass %>% 
   mutate(site = "Portal")
adjusted_portal_biomass
```

Add proportional biomass column:

```{r}
#divide biomasses for each site by the highest biomass recorded for each species
max_orpi_biomass <- orpi_biomass%>% 
  group_by(species) %>% 
  summarize(max_biomass_g = max(biomass_g))

orpi_biomass <- orpi_biomass %>% 
  left_join(max_orpi_biomass, join_by(species)) %>% 
  mutate(prop_biomass = biomass_g / max_biomass_g)
orpi_biomass

max_saguaro_biomass <- saguaro_biomass %>% 
  group_by(species) %>% 
  summarize(max_biomass_g = max(biomass_g))

saguaro_biomass <- saguaro_biomass %>% 
  left_join(max_saguaro_biomass, join_by(species)) %>% 
  mutate(prop_biomass = biomass_g / max_biomass_g)
saguaro_biomass

max_portal_biomass <- adjusted_portal_biomass %>% 
  group_by(species) %>% 
  summarize(max_biomass_g = max(biomass_g))

adjusted_portal_biomass <- adjusted_portal_biomass %>% 
  left_join(max_portal_biomass, join_by(species)) %>% 
  mutate(prop_biomass = biomass_g / max_biomass_g)
adjusted_portal_biomass
```

Bind dataframes into one:

```{r}
all_biomass <- bind_rows(saguaro_biomass, orpi_biomass, adjusted_portal_biomass)
all_biomass
```

Plotting annual biomass of each species:

```{r}
all_biomass %>% 
    filter(species %in% c("albigula", "amplus", "baileyi", "eremicus", "intermedius", "merriami", "penicillatus", "spectabilis", "torridus")) %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species, scales = "free_y") +
    labs(y = "Biomass (g)", x = "Year") +
    theme_bw()


ggsave("../figures/all_biomass.png")


all_biomass %>% 
  filter(species == "albigula") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "N. albigula Biomass (g)", x = "Year") +
    theme_bw()

all_biomass %>% 
  filter(species == "baileyi") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "C. baileyi Biomass (g)", x = "Year") +
    theme_bw()

all_biomass %>% 
  filter(species == "eremicus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "C. eremicus Biomass (g)", x = "Year") +
    theme_bw()

all_biomass %>% 
  filter(species == "merriami") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. merriami Biomass (g)", x = "Year") +
    theme_bw()

all_biomass %>% 
  filter(species == "penicillatus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "C. penicillatus Biomass (g)", x = "Year") +
    theme_bw()

all_biomass %>% 
  filter(species == "torridus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "O. torridus Biomass (g)", x = "Year") +
    theme_bw()
```

#### By community:

Adjusting Portal's biomass for multiple surveys per year:

```{r}
#loading in data
portal_total_biomass <- read_csv("../data_clean/portal_total_biomass.csv")

adjusted_portal_total_biomass <- portal_total_biomass %>% 
  left_join(portal_periods_annual) %>% 
  mutate(adjusted_biomass_g = biomass_g / n_periods) %>% 
  select(1, 4) %>% 
  rename(biomass_g = adjusted_biomass_g)
adjusted_portal_total_biomass
```

Add site column to all dataframes:

```{r}
#loading in the rest of the data
saguaro_total_biomass <- read_csv("../data_clean/saguaro_total_biomass.csv")
orpi_total_biomass <- read_csv("../data_clean/orpi_total_biomass.csv")

#adding site column to each dataframe
saguaro_total_biomass <- saguaro_total_biomass %>% 
  mutate(site = "Saguaro")
saguaro_total_biomass

orpi_total_biomass <- orpi_total_biomass %>% 
   mutate(site = "Organ")
orpi_total_biomass

adjusted_portal_total_biomass <- adjusted_portal_total_biomass %>% 
   mutate(site = "Portal")
adjusted_portal_total_biomass
```

Add proportional biomass column:

```{r}
#divide biomasses for each site by the highest annual biomass recorded
orpi_total_biomass <- orpi_total_biomass %>% 
  mutate(prop_biomass = biomass_g / max(biomass_g))
orpi_total_biomass

saguaro_total_biomass <- saguaro_total_biomass %>% 
  mutate(prop_biomass = biomass_g / max(biomass_g))
saguaro_total_biomass

adjusted_portal_total_biomass <- adjusted_portal_total_biomass %>% 
  mutate(prop_biomass = biomass_g / max(biomass_g))
adjusted_portal_total_biomass
```

Bind dataframes into one:

```{r}
all_total_biomass <- bind_rows(saguaro_total_biomass, orpi_total_biomass, adjusted_portal_total_biomass)
```

Plotting community biomass:

```{r}
#facet_wrapped
ggplot(all_total_biomass, aes(x = year, y = prop_biomass)) +
  geom_point() +
  geom_line() +
  facet_wrap(~site, scales = "free_y", ncol = 1) +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
ggsave("../figures/all_total_biomass.png")

#by color
ggplot(all_total_biomass, aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
```

```{r}
#just saguaro and orpi NOT PROPORTIONS
all_total_biomass %>% 
  filter(site != "Portal") %>% 
  ggplot(., aes(x = year, y = biomass_g, color = site)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year", title = "Saguaro National Park and Organ Pipe National Monument") +
  theme_bw()
ggsave("../figures/saguaro_orpi_total_biomass.png")
```

### Abundance Data

Loading in datasets:

```{r}
saguaro_abund <- read_csv("../data_clean/saguaro_abund_trapnights.csv")
orpi_abund <- read_csv("../data_clean/orpi_abund_trapnights.csv")
portal_abund <- read_csv("../data_clean/portal_abund_trapnights.csv")

#removing superfluous columns
saguaro_abund <- saguaro_abund %>% 
  select(1, 2, 6)
saguaro_abund

orpi_abund <- orpi_abund %>% 
  select(1, 2, 5)
orpi_abund

portal_abund <- portal_abund %>% 
  select(1, 2, 5)
portal_abund

#adding site column to each dataframe
saguaro_abund <- saguaro_abund %>% 
  mutate(site = "Saguaro")
saguaro_abund

orpi_abund <- orpi_abund %>% 
   mutate(site = "Organ Pipe")
orpi_abund

portal_abund <- portal_abund %>% 
   mutate(site = "Portal")
portal_abund
```

Modifying Portal's species column to match:

```{r}
portal_abund <- portal_abund %>% 
  left_join(adjusted_portal_species, join_by(species)) %>% 
  select(1, 5, 3, 4) %>% 
  rename(species = species_name)
portal_abund
```

Bind dataframes into one

```{r}
all_abund <- bind_rows(saguaro_abund, orpi_abund, portal_abund)
all_abund
```

Plotting abundance of each species:

```{r}
all_abund %>% 
  filter(species %in% c("albigula", "amplus", "baileyi", "eremicus", "intermedius", "merriami", "penicillatus", "spectabilis", "torridus")) %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species, scales = "free_y") +
    labs(y = "Abundance per Trapnight", x = "Year") +
    scale_y_log10() +
    theme_bw()
ggsave("../figures/all_abund.png")

all_abund %>% 
  filter(species == "baileyi") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "C. baileyi Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund %>% 
  filter(species == "torridus") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "O. torridus Abundance per Trapnight", x = "Year") +
    theme_bw()
```

# Comparing Chihuahuan Sites with Portal

### Biomass Data

#### By species

Add site column to dataframes:

```{r}
#Read in data
sevilleta_biomass <- read_csv("../data_clean/sevilleta_biomass.csv")
jornada_biomass <- read_csv("../data_clean/jornada_biomass.csv")

sevilleta_biomass <- sevilleta_biomass %>% 
  mutate(site = "Sevilleta")

jornada_biomass <- jornada_biomass %>% 
  mutate(site = "Jornada")
```

Add proportional biomass column:

```{r}
#divide biomasses for each site by the highest biomass recorded for each species
max_sevilleta_biomass <- sevilleta_biomass%>% 
  group_by(species) %>% 
  summarize(max_biomass_g = max(biomass_g))

sevilleta_biomass <- sevilleta_biomass %>% 
  left_join(max_sevilleta_biomass, join_by(species)) %>% 
  mutate(prop_biomass = biomass_g / max_biomass_g)
sevilleta_biomass

max_jornada_biomass <- jornada_biomass%>% 
  group_by(species) %>% 
  summarize(max_biomass_g = max(biomass_g))

jornada_biomass <- jornada_biomass %>% 
  left_join(max_jornada_biomass, join_by(species)) %>% 
  mutate(prop_biomass = biomass_g / max_biomass_g)
jornada_biomass
```

Bind dataframes into one:

```{r}
all_biomass_2 <- bind_rows(sevilleta_biomass, jornada_biomass, adjusted_portal_biomass)
all_biomass_2
```

Plotting annual biomass of each species:

```{r}
all_biomass_2 %>% 
   filter(species %in% c("arenicola", "flavus", "hispidus", "leucodon", "leucogaster", "leucopus", "maniculatus", "megalotis", "merriami", "micropus", "ordii", "spectabilis", "spilosoma")) %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species) +
    labs(y = "Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "arenicola") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "O. arenicola Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "flavus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. flavus Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "leucogaster") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "O. leucogaster Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "leucopus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. leucopus Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "maniculatus") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. maniculatus Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "megalotis") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "R. megalotis Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "merriami") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. merriami Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "ordii") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. ordii Biomass (g)", x = "Year") +
    theme_bw()

all_biomass_2 %>% 
  filter(species == "spectabilis") %>% 
  ggplot(., aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. spectabilis Biomass (g)", x = "Year") +
    theme_bw()
```

#### By community

Add site column to all dataframes:

```{r}
#loading in the rest of the data
sevilleta_total_biomass <- read_csv("../data_clean/sevilleta_total_biomass.csv")
jornada_total_biomass <- read_csv("../data_clean/jornada_total_biomass.csv")

#adding site column to each dataframe
sevilleta_total_biomass <- sevilleta_total_biomass %>% 
  mutate(site = "Sevilleta")
sevilleta_total_biomass

jornada_total_biomass <- jornada_total_biomass %>% 
   mutate(site = "Jornada")
jornada_total_biomass
```

Add proportional biomass column:

```{r}
sevilleta_total_biomass <- sevilleta_total_biomass %>% 
  mutate(prop_biomass = biomass_g / max(biomass_g))
sevilleta_total_biomass

jornada_total_biomass <- jornada_total_biomass %>% 
  mutate(prop_biomass = biomass_g / max(biomass_g))
jornada_total_biomass
```

Bind dataframes into one:

```{r}
all_total_biomass_2 <- bind_rows(sevilleta_total_biomass, jornada_total_biomass, adjusted_portal_total_biomass)
```

Plotting community biomass:

```{r}
#facet_wrapped
ggplot(all_total_biomass_2, aes(x = year, y = prop_biomass)) +
  geom_point() +
  geom_line() +
  facet_wrap(~site, scales = "free_y", ncol = 1) +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
ggsave("../figures/all_total_biomass.png")

#by color
ggplot(all_total_biomass_2, aes(x = year, y = prop_biomass, color = site)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
```

### Abundance Data

Loading in datasets:

```{r}
sevilleta_abund <- read_csv("../data_clean/sevilleta_abund_trapnights.csv")
jornada_abund <- read_csv("../data_clean/jornada_abund_trapnights.csv")

#removing superfluous columns
sevilleta_abund <- sevilleta_abund %>% 
  select(1,2,5)
sevilleta_abund

jornada_abund <- jornada_abund %>% 
  select(1,2,5)
jornada_abund

#adding site column to each dataframe
sevilleta_abund <- sevilleta_abund %>% 
  mutate(site = "Sevilleta")
sevilleta_abund

jornada_abund <- jornada_abund %>% 
  mutate(site = "Jornada")
jornada_abund
```

Bind dataframes into one:

```{r}
all_abund_2 <- bind_rows(sevilleta_abund, jornada_abund, portal_abund)
all_abund_2
```

Plotting abundance of each species:

```{r}
all_abund_2 %>% 
  filter(species %in% c("arenicola", "flavus", "hispidus", "leucodon", "leucogaster", "leucopus", "maniculatus", "megalotis", "merriami", "micropus", "ordii", "spectabilis", "spilosoma")) %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species, scales = "free_y") +
    labs(y = "Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "flavus") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. flavus Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "leucogaster") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "O. Lecuogaster Abundance per Trapnight", x = "Year") +
    theme_bw()


all_abund_2 %>% 
  filter(species == "leucopus") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. leucopus Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "maniculatus") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "P. maniculatus Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "megalotis") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "R. megalotis Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "merriami") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. merriami Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "ordii") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. ordii Abundance per Trapnight", x = "Year") +
    theme_bw()

all_abund_2 %>% 
  filter(species == "spectabilis") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight, color = site)) +
  geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    labs(y = "D. spectabilis Abundance per Trapnight", x = "Year") +
    theme_bw()

```
