---
title: "data_comparison"
author: "Hannah Dempsey"
output: github_document
---

# Comparing Field Site Statistics

Set up:

```{r}
library(tidyverse)
library(portalr)
```

### Biomass Data

#### By species:

Adjusting Portal's biomass for multiple surveys per year:

```{r}
#Loading in data
portal_biomass <- read_csv("../data_clean/portal_biomass.csv")
portal_trapping <- read_csv("../data_clean/portal_trapping.csv")

#calculating number of periods per year
portal_periods <- portal_trapping %>% 
  group_by(year) %>% 
  summarize(periods = unique(period))
portal_periods
 
portal_periods_annual <- portal_periods %>% 
  group_by(year) %>% 
  summarize(n_periods = n())
portal_periods_annual

#dividing biomass by number of periods
adjusted_portal_biomass <- portal_biomass %>% 
  inner_join(portal_periods_annual, join_by(year)) %>% 
  mutate(adjusted_biomass_g = biomass_g / n_periods) %>% 
  select(1:2, 7) %>% 
  #renaming so all column names across sites are the same
  rename(biomass_g = adjusted_biomass_g)
adjusted_portal_biomass
```

Modify portal's species column to match:

```{r}
#load in species data
portal_species <- read_csv("../data_clean/portal_species.csv")

#distinguish between hispidus spp
adjusted_portal_species <- portal_species %>% 
  mutate(scientificname = str_replace(scientificname, "Chaetodipus hispidus", " C.hispidus")) %>% 
  mutate(scientificname = str_replace(scientificname, "Sigmodon hispidus", " S.hispidus")) %>% 
  separate(scientificname, c("genus", "species_name"), sep = " ") %>% 
  select(1, 3)
adjusted_portal_species

#join to biomass dataframe
adjusted_portal_biomass <- adjusted_portal_biomass %>% 
  left_join(adjusted_portal_species, join_by(species)) %>% 
  select(1, 4, 3) %>% 
  rename(species = species_name)
adjusted_portal_biomass
```

Adding site column to all dataframes:

```{r}
#loading in the rest of the data
saguaro_biomass <- read_csv("../data_clean/saguaro_biomass.csv")
orpi_biomass <- read_csv("../data_clean/orpi_biomass.csv")

#adding site column to each dataframe
saguaro_biomass <- saguaro_biomass %>% 
  mutate(site = "Saguaro")
saguaro_biomass

orpi_biomass <- orpi_biomass %>% 
   mutate(site = "Organ Pipe")
orpi_biomass

adjusted_portal_biomass <- adjusted_portal_biomass %>% 
   mutate(site = "Portal")
adjusted_portal_biomass
```

Bind dataframes into one:

```{r}
all_biomass <- bind_rows(saguaro_biomass, orpi_biomass, adjusted_portal_biomass)
all_biomass
```

Plotting annual biomass of each species:

```{r}
all_biomass %>% 
  filter(species %in% c("albigula", "amplus", "baileyi", "eremicus", "intermedius", "maniculatus", "merriami", "spectabilis", "torridus")) %>% 
  ggplot(., aes(x = year, y = biomass_g, color = site)) +
    geom_point(alpha = 0.5) +
    geom_line(alpha = 0.5) +
    facet_wrap(~species, scales = "free_y") +
    labs(y = "Biomass (g)", x = "Year") +
    theme_bw()
ggsave("../figures/all_biomass.png")
```

#### By community:

Adjusting Portal's biomass for multiple surveys per year:

```{r}
#loading in data
portal_total_biomass <- read_csv("../data_clean/portal_total_biomass.csv")

adjusted_portal_total_biomass <- portal_total_biomass %>% 
  left_join(portal_periods_annual) %>% 
  mutate(adjusted_biomass_g = biomass_g / n_periods) %>% 
  select(1, 4) %>% 
  rename(biomass_g = adjusted_biomass_g)
adjusted_portal_total_biomass
```

Adding site column to all dataframes:

```{r}
#loading in the rest of the data
saguaro_total_biomass <- read_csv("../data_clean/saguaro_total_biomass.csv")
orpi_total_biomass <- read_csv("../data_clean/orpi_total_biomass.csv")

#adding site column to each dataframe
saguaro_total_biomass <- saguaro_total_biomass %>% 
  mutate(site = "Saguaro")
saguaro_total_biomass

orpi_total_biomass <- orpi_total_biomass %>% 
   mutate(site = "Organ")
orpi_total_biomass

adjusted_portal_total_biomass <- adjusted_portal_total_biomass %>% 
   mutate(site = "Portal")
adjusted_portal_total_biomass
```

Bind dataframes into one:

```{r}
all_total_biomass <- bind_rows(saguaro_total_biomass, orpi_total_biomass, adjusted_portal_total_biomass)
```

Plotting community biomass:

```{r}
#facet_wrapped
ggplot(all_total_biomass, aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  facet_wrap(~site, scales = "free_y", ncol = 1) +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
ggsave("../figures/all_total_biomass.png")

#by color
ggplot(all_total_biomass, aes(x = year, y = biomass_g, color = site)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year") +
  theme_bw()
```

```{r}
#just saguaro and orpi
all_total_biomass %>% 
  filter(site != "Portal") %>% 
  ggplot(., aes(x = year, y = biomass_g, color = site)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year", title = "Saguaro National Park and Organ Pipe National Monument") +
  theme_bw()
ggsave("../figures/saguaro_orpi_total_biomass.png")
```
