---
title: "orpi_analysis"
author: "Hannah Dempsey"
output: github_document
---

# Organ Pipe National Monument Analysis

Set up:

```{r}
library(tidyverse)
orpi_clean <- read_csv("../data_clean/orpi_clean.csv")
orpi_surveys <- read_csv("../data_clean/orpi_surveys.csv")
```

### Abundance Data

Summarize the survey data by year:

```{r}
trap_nights <- orpi_surveys %>% 
  mutate(year = year(end_date),
         nights = end_date - start_date,
         trap_nights = num_traps * nights) %>% 
  group_by(year) %>% 
  summarize(trap_nights = as.numeric(sum(trap_nights)))
trap_nights
```

Calculate the annual abundance for each species per trapnight:

```{r}
rodents_per_trapnight <- orpi_clean %>% 
  mutate(year = year(end_date)) %>%  
  group_by(year, species) %>% 
  summarize(abundance = sum(recapture == 0)) %>% 
  inner_join(trap_nights, join_by(year)) %>% 
  mutate(abund_per_trapnight = abundance / trap_nights)
rodents_per_trapnight

#save dataframe for comparison
write_csv(rodents_per_trapnight, "../data_clean/orpi_abund_trapnights.csv")
```

Plotting abundance per trapnight:

```{r}
#only one data point for maniculatus, so I'm removing it from the graph
orpi_abundance <- rodents_per_trapnight %>% 
  filter(species != "maniculatus") %>% 
  ggplot(., aes(x = year, y = abund_per_trapnight)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ species) +
  labs(y = "Rodent Abundance per Trapnight (g)", x = "Year", title = "Organ Pipe National Monument") +
  theme_bw()
orpi_abundance
ggsave("../figures/orpi_abundance.png")
```

### Biomass Data

Calculating average mass for each species:

```{r}
rodent_mass <- orpi_clean %>% 
  group_by(species) %>% 
  summarize(avg_mass_g = mean(weight, na.rm = TRUE))
rodent_mass
```

Calculate the annual biomass of each species:

```{r}
biomass_annual <- rodents_per_trapnight %>% 
  inner_join(rodent_mass, join_by(species)) %>% 
  group_by(year, species) %>% 
  summarize(biomass_g = abund_per_trapnight * avg_mass_g)
biomass_annual

#save dataframe for comparison
write_csv(biomass_annual, "../data_clean/orpi_biomass.csv")
```

Plotting annual biomass of each species:

```{r}
#only one data point for maniculatus, so I'm removing it from the graph
orpi_biomass <- biomass_annual %>% 
  filter(species != "maniculatus") %>% 
  ggplot(., aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species) +
  labs(y = "Biomass (g)", x = "Year", title = "Organ Pipe National Monument") +
  theme_bw()
orpi_biomass
ggsave("../figures/orpi_biomass.png")
```

Calculating annual community biomass:

```{r}
total_biomass_annual <- biomass_annual %>% 
  group_by(year) %>% 
  summarize(biomass_g = sum(biomass_g))
total_biomass_annual

#save dataframe for comparison
write_csv(total_biomass_annual, "../data_clean/orpi_total_biomass.csv")
```

Plotting annual community biomass:

```{r}
orpi_total_biomass <- ggplot(total_biomass_annual, aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year", title = "Organ Pipe National Monument") +
  theme_bw()
orpi_total_biomass
ggsave("../figures/orpi_total_biomass.png")
```
