---
title: "portal_analysis"
author: "Hannah Dempsey"
output: github_document
---

# Portal Analysis

Set up:

```{r}
library(tidyverse)
library(portalr)
portal_abund <- read_csv("../data_clean/portal_abund.csv")
```

### Abundance Data

Calculate the annual abundance for each species per trapnight:

```{r}
rodents_per_trapnight <- portal_abund %>% 
  group_by(year, species) %>% 
  summarize(abundance = sum(abundance, na.rm = T),
            effort = sum(effort, na.rm = T)) %>% 
  mutate(abund_per_trapnight = abundance / effort)
rodents_per_trapnight

#save dataframe for comparison
write_csv(rodents_per_trapnight, "../data_clean/portal_abund_trapnights.csv")
```

Plotting abundance per trapnight:

```{r}
portal_abundance <- ggplot(rodents_per_trapnight, aes(x = year, y = abund_per_trapnight)) +
  geom_point(size = 0.5) +
  geom_line() +
  facet_wrap(~ species) +
  labs(y = "Rodent Abundance per Trapnight", x = "Year", title = "Portal Project") +
  theme_bw()
portal_abundance
ggsave("../figures/portal_abundance.png")
```

### Biomass Data

Calculate the annual biomass of each species:

```{r}
biomass_annual <- portal_abund %>% 
  select(species, avg_weight_g) %>% 
  right_join(rodents_per_trapnight, join_by(species)) %>% 
  select(3, 1:2, 6) %>% 
  mutate(biomass_g = abund_per_trapnight * avg_weight_g)
biomass_annual

#save dataframe for comparison
write_csv(biomass_annual, "../data_clean/portal_biomass.csv")
```

Plotting annual biomass of each species:

```{r}
portal_biomass <- ggplot(biomass_annual, aes(x = year, y = biomass_g)) +
  geom_point(size = 0.5) +
  geom_line() +
  facet_wrap(~species) +
  labs(y = "Biomass (g)", x = "Year", title = "Portal Project") +
  theme_bw()
portal_biomass
ggsave("../figures/portal_biomass.png")
```

Calculating annual community biomass:

```{r}
total_biomass_annual <- biomass_annual %>% 
  group_by(year) %>% 
  summarize(biomass_g = sum(biomass_g))
total_biomass_annual

#save dataframe for comparison
write_csv(total_biomass_annual, "../data_clean/portal_total_biomass.csv")
```

Plotting annual community biomass:

```{r}
portal_total_biomass <- ggplot(total_biomass_annual, aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year", title = "Portal Project") +
  theme_bw()
portal_total_biomass
ggsave("../figures/portal_total_biomass.png")
```

### Weather Analysis

Read in the weather data

```{r}
weather_data <- weather("Monthly")
```

Summarize annual statistics

```{r}
weather_avg <- weather_data %>% 
  group_by(year) %>% 
  summarize(avg_mintemp = mean(mintemp, na.rm = T),
            avg_maxtemp = mean(maxtemp, na.rm = T),
            avg_meantemp = mean(meantemp, na.rm = T),
            avg_precip = mean(precipitation, na.rm = T))
weather_avg
```

Graphing weather:

```{r}
ggplot(weather_avg, aes(x = year, y = avg_precip)) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Average Precipitation", title = "Portal Project") +
  theme_bw()

ggplot(weather_avg, aes(x = year, y = avg_meantemp)) +
  geom_point() +
  geom_line() +
  labs(x = "Year", y = "Average Temperature", title = "Portal Project") +
  theme_bw()
```
