---
title: "saguaro_analysis"
author: "Hannah Dempsey"
output: github_document
---

# Saguaro National Park Analysis

Set up:

```{r}
library(tidyverse)
saguaro_clean <- read_csv("../data_clean/saguaro_clean.csv")
```

### Abundance Data

Summarize the data by year:

```{r}
survey_nights <- saguaro_clean %>% 
  group_by(year) %>% 
  distinct(trap_check_date) %>% 
  summarize(nights = n())
survey_nights
```

Calculate the annual abundance for each species per trapnight:

```{r}
#148 traps used per night
rodents_per_trapnight <- saguaro_clean %>% 
  filter(recapture == "N") %>% 
  group_by(year, species) %>% 
  summarize(abundance = n()) %>% 
  left_join(survey_nights) %>% 
  mutate(trap_nights = nights * 148,
         abund_per_trapnight = abundance / trap_nights)
rodents_per_trapnight

#save dataframe for comparison
write_csv(rodents_per_trapnight, "../data_clean/saguaro_abund_trapnights.csv")
```

Plotting abundance per trapnight:

```{r}
saguaro_abundance <- ggplot(rodents_per_trapnight, aes(x = year, y = abund_per_trapnight)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ species) +
  labs(y = "Rodent Abundance per Trapnight", x = "Year", title = "Saguaro National Park") +
  theme_bw()
saguaro_abundance
ggsave("../figures/saguaro_abundance.png")
```

### Biomass Data

Calculating average mass for each species:

```{r}
rodent_mass <- saguaro_clean %>% 
  group_by(species) %>% 
  summarize(avg_mass_g = mean(mass_g, na.rm = TRUE))
rodent_mass
```

Calculate annual biomass of each species:

```{r}
biomass_annual <- rodents_per_trapnight %>% 
  inner_join(rodent_mass, join_by(species)) %>% 
  group_by(year, species) %>% 
  summarize(biomass_g = abund_per_trapnight * avg_mass_g)
biomass_annual

#divide biomass by trapnights
trap_nights_per_year <- rodents_per_trapnight %>% 
  group_by(year) %>% 
  summarize(trap_nights = unique(trap_nights))
trap_nights_per_year

biomass_annual <- biomass_annual %>% 
  left_join(trap_nights_per_year, join_by(year)) %>% 
  mutate(biomass_g = biomass_g / trap_nights)
biomass_annual
  
#save dataframe for comparison
write_csv(biomass_annual, "../data_clean/saguaro_biomass.csv")
```

Plotting annual biomass of each species:

```{r}
saguaro_biomass <- ggplot(biomass_annual, aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species) +
  labs(y = "Biomass (g)", x = "Year", title = "Saguaro National Park") +
  theme_bw()
saguaro_biomass
ggsave("../figures/saguaro_biomass.png")
```

Calculating annual community biomass:

```{r}
total_biomass_annual <- biomass_annual %>% 
  group_by(year) %>% 
  summarize(biomass_g = sum(biomass_g))
total_biomass_annual

#save dataframe for comparison
write_csv(total_biomass_annual, "../data_clean/saguaro_total_biomass.csv")
```

Plotting annual community biomass:

```{r}
saguaro_total_biomass <- ggplot(total_biomass_annual, aes(x = year, y = biomass_g)) +
  geom_point() +
  geom_line() +
  labs(y = "Total Biomass (g)", x = "Year", title = "Saguaro National Park") +
  theme_bw()
saguaro_total_biomass
ggsave("../figures/saguaro_total_biomass.png")
```
